$\pi$ with leftovers: \\ a mechanisation in Agda

### 2020-04-29 paper/paper.tex
Pass on syntax

The idea de Bruijn had \cite{} was to use a natural number $n$ (aka \emph{index}) to refer to the variable introduced $n$ binders ago --- and binders no longer introduce names.
\begin{example}
The \picalc{} term $(\textbf{new} \; x) (x ( y ) . y \langle \texttt{"Hello world!"} \rangle . \mathbf{0} \parallel (\textbf{new} \; y) (x \langle y \rangle . y ( z ) .\mathbf{0}))$ would translate to $(\textbf{new}) (0 () . 0 \langle \texttt{"Hello world"} \rangle . \mathbf{0} \parallel (\textbf{new}) (1 \langle 0 \rangle . 0 () . \mathbf{0}))$.
\end{example}
That is, terms at different \emph{depths} use different indices to refer to the same binding.
Humans often find this confusing.
There is however no reason not to annotate these indices with the original names bestowed by humans.
Machines can then manipulate references mechanically while still using names to present terms to humans.

A variable occurring under $n$ abstractions has $n$ things to refer to.
References outside of that range have no associated meaning.
It is useful to rule out these nonsensical terms syntactically.
In Figure \ref{fig:var} we do so by introducing the indexed family of types $\Var_n$: for all naturals $n$, the type $\Var_n$ has $n$ distinct elements.

@@

Using type-level de Bruijn indices makes our syntax well-scoped by construction, and the semantics of our language can now be defined on the totality of the syntax.
User-friendliness can still be recovered through a function that converts processes with names into (possibly) processes with indices.
This function keeps track of what index is associated with what name, and traverses the process recursively, taking note of new binders and substituting variable references.
If the process is ill-scoped, the function returns nothing.
To present things back to the user, we print indices as their named annotations.

>>>>

The idea de Bruijn had \cite{} was to use a natural number $n$ (aka \emph{index}) to refer to the variable introduced $n$ binders ago --- and binders no longer introduce names.
\begin{example}
The \picalc{} term $(\textbf{new} \; x) (x ( y ) . y \langle \texttt{"Hello world!"} \rangle . \mathbf{0} \parallel (\textbf{new} \; y) (x \langle y \rangle . y ( z ) .\mathbf{0}))$ would translate to $(\textbf{new}) (0 () . 0 \langle \texttt{"Hello world"} \rangle . \mathbf{0} \parallel (\textbf{new}) (1 \langle 0 \rangle . 0 () . \mathbf{0}))$.
\end{example}
That is, terms at different \emph{depths} use different indices to refer to the same binding.
Humans may find this confusing.
There is however no reason not to annotate these indices with the original names bestowed upon them by humans.
Machines can then manipulate references mechanically whilst using names to present terms to humans.

A variable occurring under $n$ abstractions has $n$ possible variables to refer to.
References outside of that range are meaningless.
It is useful to rule out these nonsensical terms syntactically.
In Figure \ref{fig:var} we do so by introducing the indexed family of types $\Var_n$: for all naturals $n$, the type $\Var_n$ has $n$ distinct elements.

@@

Using type-level de Bruijn indices makes our syntax well-scoped by construction --- the semantics of our language can now be defined on the totality of the syntax.
User-friendliness can still be recovered through a function that converts processes with named variables into (possibly) processes with indexed variables.
This function recursively traverses processes, keeping track of what name is associated with each index.
If the process is ill-scoped, the function returns nothing.
To present things to the user, we substitute indices by their related names.
